Project-I by Group MexicoCity

### 2014-11-03 report/paper.tex
update

\textcolor{red}{Formula? Not even sure this section is useful}

\section{Data observation}
\label{sec:datadescr}
We have two data set. One for regression and one for classification. \begin{description}
\item[Regression] Consists of output variables $\mathbf{y}$ and input variables $\mathbf{X}$. The number of examples is $N = 1400$ and each input $\mathbf{x_n}$ has dimensionality $D = 48$. The first $34$ are real valued and the last $14$ are categorical, included $5$ that are binaries.
\item[Classification] Consists of output variables $\mathbf{y}$ and input variables $\mathbf{X}$. The number of examples is $N = 1500$ and each input $\mathbf{x_n}$ has dimensionality $D = 35$. \textcolor{red}{check hist}

@@

\textcolor{red}{need to move that somewhere}Note that all the gradient descent methods used the eps threshold (machine dependant) in Matlab in the gradient descent function, to allow for a maximum number of iterations. This might make the gradient descent process slow though, so feel free to set another number if needed. We changed it later to 1/1000, so the numbers for the test Errors might be different depending on this parameter when you run the code (the global trends are the same though), since the number of iterations will be different.

@@

We tested taking the squares of the 18th and the 34th variable individually and a curious thing happened: when we removed the 18th variable (squared) the test error increased to to 0.3881, but when we removed the 34th variable (squared) the test error increased much more significantly to 0.4987.

\subsection{Ridge Regression}
\label{sec:regridgreg}
By visually inspecting the Pearson correlation graphs, a good strategy to spot well correlated input variables was to take the intersection of the two methods (we’re only considering ). To do that, we tried multiplying the two coefficients (for each variable) and taking the mean. Both strategies clearly showed the 18th and 34th variables are best correlated. We can limit our variable transform to these two (we only have 1400 data points, so with a dimension 48 we’re already lacking data, by introducing spurious transforms we would be adding fuel to the fire and causing overfitting.)\\
We have also done a cross validation using seeds. The result is shown in Figure \ref{fig:ridreggcv}.
\begin{figure}
\centering
\includegraphics[width=1\textwidth]{images/ridgeRegCV.eps}
\caption{Cross validation : Ridge regression with different lambda values }
\label{fig:ridreggcv}
\end{figure}

\subsection{Logistic Regression}
We tried to use simple logistic regression using all the variable, the estimated RMSE test error was 0.27099. Also we tried using sqrt on the 11th and 24th variables which actually improves our test error to 0.26779. We can also use squares of the variables instead of the square root and applying this alone the test error decreased to 0.26715. Adding both transformations got us a decrease to a test error of 0.26474. We also tried to remove all categorical variables as we did in regression, but the test error increased to 0.3396, so their information was needed and cannot be ignored.
\subsection{Feature transformations}
By isolating the 18th and 34th input variables, and producing polynomial transforms of them using the mypoly function, and then measuring their Pearson correlation coefficient with the output, we could see that the cube and the power 5 of the 18th variable, and powers up to 6 for the 34th variable are correlated with the output. When adding these feature transformations, we could lower the test error to 0.3307, which is quite lower now than the starting test error of 0.3805.\\

For classification, the rank of $\mathbf{\tilde{X}}$ is 35 plus the one vector. Given the data description from section \ref{sec:datadescr}, this tells us that $\mathbf{\tilde{X}}$ is not ill-conditioned.\\

\subsection{Ridge Regression}
For the classification, we repeated the procedure describe in section \ref{sec:regridgreg}. We found that the 11th and the 24th variables are best correlated with the output, therefore we used feature transforms on those. \textcolor{red}{what transformation? sqrt etc but only on these variables? }

\subsection{Penalized logistic regression}
We could improve the test error by cross-validation. The Figure \ref{fig:crossval} shows the three different test error RMSE, 0-1 loss and logloss. We used different numbers of seed to have a better prediction for our test error.
\begin{figure}
\centering
\subfigure[0-1 loss]{\includegraphics[width=1\textwidth]{images/penLogRegCrossVall01Loss.eps}}
\subfigure[rmse]{\includegraphics[width=1\textwidth]{images/penLogRegrmse.eps}}
\subfigure[log loss]{\includegraphics[width=1\textwidth]{images/penLogRegLogloss.eps}}
\caption{Cross-validation : penalized logistic regression with different lambda values}
\label{fig:crossval}
\end{figure} 
\subsection{Feature transformations}

We tried dummy variables for all regression categorical ones. We could only find a significant linear correlation (using Spearman) for the 2nd class of the variable 44, 2nd and 4th class of the 47th.\\

@@

\section{Summary}
As we have seen, the cross validation helps to have a better test error.\\

We obtained these RMSE for regression.
\begin{description}
\item[leastSquaresGD] We achieved a RMSE of 0.6852
\item[leastSquares] We achieved a RMSE of 0.3317
\item[RidgeRegression] We achieved a RMSE of 0.3073
\end{description}
For the regression, we saw that the best method was Ridge Regression. The best parameter was $\lambda =  39.0694$\textcolor{red}{?} and we achieved a test error of 0.3073.

For the classification, the best method was penalized logistic regression. We obtained different results depending the type of error we computed. \begin{description}
\item[RMSE] We achieved a test error of 0.2644.
\item[01 loss]We achieved a test error of 0.09486.
\item[log loss] We achieved a test error of 0.25149.
\end{description}
With logistic regression we obtained different results depending the type of error we computed. \begin{description}
\item[rmse] We achieved a test error of 0.2644.
\item[01 loss]We achieved a test error of 0.09486.
\item[log loss] We achieved a test error of 0.2637.
\end{description}
We see that there is only a small improvement with the log loss error.

>>>>

\section{Data observation}
\label{sec:datadescr}
We have two data set. One for regression and one for classification. \begin{description}
\item[Regression] Consists of output variables $\mathbf{y}$ and input variables $\mathbf{X}$. The number of examples is $N = 1400$ and each input $\mathbf{x_n}$ has dimensionality $D = 48$. The first $34$ are real valued and the last $14$ are categorical, included $5$ that are binaries.
\item[Classification] Consists of output variables $\mathbf{y}$ and input variables $\mathbf{X}$. The number of examples is $N = 1500$ and each input $\mathbf{x_n}$ has dimensionality $D = 35$.

@@

Note that all the gradient descent methods used the eps threshold (machine dependant) in Matlab in the gradient descent function, to allow for a maximum number of iterations. This might make the gradient descent process slow though, so feel free to set another number if needed. We changed it later to 1/1000, so the numbers for the test Errors might be different depending on this parameter when you run the code (the global trends are the same though), since the number of iterations will be different.

@@

We tested taking the squares of the 18th and the 34th variable individually and a curious thing happened: when we removed the 18th variable (squared) the test error increased to to 0.3881, but when we removed the 34th variable (squared) the test error increased much more significantly to 0.4987.\\

\label{sec:regridgreg}
By visually inspecting the Pearson correlation graphs, a good strategy to spot well correlated input variables was to take the intersection of the two methods (we’re only considering ). To do that, we tried multiplying the two coefficients (for each variable) and taking the mean. Both strategies clearly showed the 18th and 34th variables are best correlated. We can limit our variable transform to these two (we only have 1400 data points, so with a dimension 48 we’re already lacking data, by introducing spurious transforms we would be adding fuel to the fire and causing overfitting.)\\

\subsection{Ridge Regression}
After applying least squares, we have also done a cross validation using seeds. The result is shown in Figure \ref{fig:ridreggcv}.
\begin{figure}
\centering
\includegraphics[width=1\textwidth]{images/ridgeRegCV.eps}
\caption{Cross validation : Ridge regression with different lambda values }
\label{fig:ridreggcv}
\end{figure}
We realized that the RMSE decreased for a parameter $\lambda = 39.0694$ to a RMSE of $0.3073$. We can see that on the Figure \ref{fig:ridreggcv}.

\subsection{Feature transformations}
We tried dummy variables for all regression categorical ones. We could only find a significant linear correlation (using Spearman) for the 2nd class of the variable 44, 2nd and 4th class of the 47th.\\

@@

By isolating the 18th and 34th input variables, and producing polynomial transforms of them using the mypoly function, and then measuring their Pearson correlation coefficient with the output, we could see that the cube and the power 5 of the 18th variable, and powers up to 6 for the 34th variable are correlated with the output. When adding these feature transformations, we could lower the test error to 0.3307, which is quite lower now than the starting test error of 0.3805.\\

For classification, the rank of $\mathbf{\tilde{X}}$ is 35 plus the one vector. Given the data description from section \ref{sec:datadescr}, this tells us that $\mathbf{\tilde{X}}$ is not ill-conditioned.\\ The range of $\mathbf{y}$ is between $-1$ and $1$. We also modified the values $-1$ from $\mathbf{y}$ to 0 to be consistent with the book.

\subsection{Logistic Regression}
For the classification, we repeated the procedure describe in section \ref{sec:regridgreg}. We found that the 11th and the 24th variables are best correlated with the output, therefore we used feature transforms on those as described in section \ref{sec:ftransclass}.

\subsection{Feature transformations}
\label{sec:ftransclass}
We tried to use simple logistic regression using all the variable, the estimated RMSE test error was 0.27099. Also we tried using sqrt on the 11th and 24th variables which actually improves our test error to 0.26779. We can also use squares of the variables instead of the square root and applying this alone the test error decreased to 0.26715. Adding both transformations got us a decrease to a test error of 0.26474. We also tried to remove all categorical variables as we did in regression, but the test error increased to 0.3396, so their information was needed and cannot be ignored.\\

\subsection{Penalized logistic regression}
We could improve the test error by cross-validation. The Figure \ref{fig:crossval} shows the three different test error RMSE, 0-1 loss and logloss. We used different numbers of seed to have a better prediction for our test error.
\begin{figure}
\centering
\subfigure[0-1 loss]{\includegraphics[width=1\textwidth]{images/penLogRegCrossVall01Loss.eps}}
\subfigure[rmse]{\includegraphics[width=1\textwidth]{images/penLogRegrmse.eps}}
\subfigure[log loss]{\includegraphics[width=1\textwidth]{images/penLogRegLogloss.eps}}
\caption{Cross-validation : penalized logistic regression with different lambda values}
\label{fig:crossval}
\end{figure} 
\section{Summary}
As we have seen, the cross validation helps to have a better test error.\\

We obtained these RMSE for regression.
\begin{description}
\item[leastSquaresGD] We achieved a RMSE of 0.6852
\item[leastSquares] We achieved a RMSE of 0.3317
\item[RidgeRegression] We achieved a RMSE of 0.3073
\end{description}
For the regression, we saw that the best method was Ridge Regression. The best parameter was $\lambda =  39.0694$ and we achieved a test error of 0.3073.

For the classification, the best method was penalized logistic regression. We obtained different results depending the type of error we computed. \begin{description}
\item[RMSE] We achieved a test error of 0.2644.
\item[01 loss]We achieved a test error of 0.09486.
\item[log loss] We achieved a test error of 0.25149.
\end{description}
With logistic regression we obtained different results depending the type of error we computed. \begin{description}
\item[rmse] We achieved a test error of 0.2644.
\item[01 loss]We achieved a test error of 0.09486.
\item[log loss] We achieved a test error of 0.2637.
\end{description}
We see that there is only a small improvement with the log loss error.\\
The best error we achieved is through the logistic regression method and using the 0-1 loss error.
